---
title: "Tools for Interpreting Machine Learning Models"
subtitle: "Communities and Crime Example"
author: "Christoph Kern"
output: html_notebook
---

## Setup

```{r results='hide', message=FALSE, warning=FALSE}
library(tidyverse)
library(caret)
library(ranger)
library(iml)
```

## Data

This notebook uses the Communities and Crime data from the UCI ML repository.

Source: https://archive.ics.uci.edu/ml/datasets/Communities+and+Crime+Unnormalized#

With variable names: https://github.com/Starignus/AppliedML_Python_Coursera/blob/master/CommViolPredUnnormalizedData.txt

```{r}
crime <- read.table("CommViolPredUnnormalizedData.txt", 
                    header = TRUE, 
                    sep = ",", 
                    quote = "\"", 
                    dec = ".", 
                    fill = TRUE, 
                    comment.char = "",
                    na.strings = "?",
                    stringsAsFactors = default.stringsAsFactors())
```

Drop some variables.

```{r}
crime <- select(crime, -c(communityname, 
                          state, 
                          countyCode, 
                          communityCode, 
                          fold,
                          murders, murdPerPop,
                          rapes, rapesPerPop,
                          robberies, robbbPerPop,
                          assaults, assaultPerPop,
                          burglaries, burglPerPop,
                          larcenies, larcPerPop,
                          autoTheft, autoTheftPerPop,
                          arsons, arsonsPerPop,
                          nonViolPerPop))
```

Delete variables with many missings.

```{r}
crime <- select(crime, -contains("Lemas"))
crime <- select(crime, -contains("Polic"))
crime <- select(crime, -contains("Offic"))
crime <- select(crime, -contains("race"))
crime <- select(crime, -c(RacialMatchCommPol, NumKindsDrugsSeiz))
```

Apply listwise deletion of missings in the remaining data.

```{r}
crime <- drop_na(crime)
```

Plot and summarize the outcome variable.

```{r}
summary(crime$ViolentCrimesPerPop)

ggplot(crime, aes(ViolentCrimesPerPop)) +
  geom_histogram(aes(y = stat(density)), binwidth = 50) +
  stat_function(
    fun = dnorm, 
    args = list(mean = mean(crime$ViolentCrimesPerPop), 
                sd = sd(crime$ViolentCrimesPerPop)), 
    col = 'red'
    )
```

Split the data into a training and a test set.

```{r}
set.seed(49043)
inTrain <- createDataPartition(crime$ViolentCrimesPerPop, 
                               p = .8, 
                               list = FALSE, 
                               times = 1)

crime_train <- crime[inTrain,]
crime_test <- crime[-inTrain,]
```

## Random Forest and Extra-Trees

Set up the evaluation method.

```{r}
ctrl  <- trainControl(method = "cv",
                      number = 10,
                      verboseIter = TRUE)
```

Specify the tuning grid.

```{r}
cols <- ncol(crime_train)

grid <- expand.grid(mtry = c(round(sqrt(cols)),
                             round(log(cols))),
                    splitrule = c("variance", "extratrees"),
                    min.node.size = c(5, 10))
grid
```

Run models with `caret`.

```{r}
set.seed(65948)
rf <- train(ViolentCrimesPerPop ~ .,
            data = crime_train,
            method = "ranger",
            trControl = ctrl,
            tuneGrid = grid)
```

Print results of the tuning process.

```{r}
rf
```

## IML

Create the `Predictor()` container. 

```{r}
X <- crime_train[which(names(crime_train) != "ViolentCrimesPerPop")]
predictor <- Predictor$new(rf, data = X, y = crime_train$ViolentCrimesPerPop)
```

### Feature importances

```{r}
set.seed(92943)
imp <- FeatureImp$new(predictor, loss = "mse", n.repetitions = 10)
```

```{r}
res <- imp$results
res$feature = factor(res$feature, levels = res$feature[order(res$importance)])

res %>%
  top_n(10, importance) %>%
  ggplot(., aes(y = feature, x = importance)) +  
  geom_segment(aes(y = feature, yend = feature, 
                   x = importance.05, xend = importance.95), 
               size = 1.25, color = "darkslategrey") +
        geom_point(size = 2.5) +
        scale_x_continuous("Feature Importance (loss: mse)", 
                           expand = c(0, 0), limits = c(1, 2)) + 
        scale_y_discrete("") +
  theme_grey(base_size = 15)

ggsave("varimp.png", width = 6, height = 6)
```

### Feature interaction

```{r}
set.seed(92943)
interact <- Interaction$new(predictor, grid.size = 10)
```

```{r}
res <- interact$results
res$.feature = factor(res$.feature, levels = res$.feature[order(res$.interaction)])

res %>%
  top_n(10, .interaction) %>%
  ggplot(aes(y = .feature, x = .interaction)) + geom_point() +
  geom_segment(aes(yend = .feature, x = 0, xend = .interaction)) +
  scale_x_continuous("Overall interaction strength") +
  scale_y_discrete("") +
  theme_grey(base_size = 15)

ggsave("interact.png", width = 6, height = 6)
```

### PDP and ICE plots

```{r}
ice <- FeatureEffect$new(predictor, method = "pdp+ice", 
                         feature = "PctKidsBornNeverMar",
                         center.at = min(crime_train$PctKidsBornNeverMar))
```

```{r}
plot(ice) +
  theme_grey(base_size = 15)

ggsave("ice.png", width = 6, height = 6)
```

### ALE plots

```{r}
ale <- FeatureEffect$new(predictor, method = "ale", feature = "PctKidsBornNeverMar")
```

```{r, warning = FALSE, message = FALSE}
plot(ale) + 
  scale_y_continuous("ALE of y") +
  theme_grey(base_size = 15)

ggsave("ale.png", width = 6, height = 6)
```

### Global surrogate model

```{r}
set.seed(92943)
tree <- TreeSurrogate$new(predictor, maxdepth = 2)
```

```{r}
plot(tree) +
  theme_grey(base_size = 15)

ggsave("tree.png", width = 6, height = 6)
```

### Local surrogate model via LIME

```{r, warning = FALSE, message = FALSE}
set.seed(92943)
example <- sample_n(crime_train, 1)
example
predict(rf, newdata = example)

lime <- LocalModel$new(predictor, x.interest = example, k = 3)
lime$results
#lime$model$beta[1:97,17]
#lime$model$a0[17]
```

```{r}
plot(lime) +
  scale_x_discrete("") +
  theme_grey(base_size = 15) +
  theme(plot.title = element_text(size = 15))

ggsave("lime.png", width = 6, height = 6)
```

## References

* https://cran.r-project.org/web/packages/iml/index.html